---
title: "Predicting Urban Growth"
author: "Austin, TX"
date: "Olivia Scalora & Hasa Reddy"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }

</style>

# 1. Introduction

```{r load_packages, message=FALSE, warning=FALSE, results = "hide"}
# install.packages("RCurl")
library(tidyverse)
library(sf)
library(raster)
library(knitr)
library(kableExtra)
library(tidycensus)
library(tigris)
library(FNN)
#library(QuantPsyc) # JE Note: in R 4.1, QuantPsyc package not available.
library(caret)
library(yardstick)
library(pscl)
library(plotROC) 
library(ggrepel)
library(pROC)
library(grid)
library(gridExtra)
library(viridis)
library(igraph)
library(RCurl)

options(scipen = 100)

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

palette2 <- c("#41b6c4","#253494")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#ffffcc","#a1dab4","#41b6c4","#2c7fb8","#253494")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
```

```{r, warning = FALSE, message = FALSE}
#this function converts a column in to quintiles. It is used for mapping.
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }
```

# 2. Data Wrangling & Feature Engineering

## 2.2. Land Cover Change Data

```{r load_data, warning = FALSE, message = FALSE, results = "hide"}
#LC Change was pre processed in ArcGIS using 2008 and 2019 Land cover data 
# Raster Calculator was used to retrieve raster cells changed from undeveloped to developed between 2008 and 2019
lc_change = raster("~/GitHub/Urban_Growth_ATX/Austin_Data/urbn_grwth_08_19.tif")

#austinMSA is the extent of the study area (Austing Metropolitan Statistical Area)
studyAreaCounties<- c("Williamson", "Travis","Bastrop", "Caldwell", "Hays")
austinMSA <- 
  st_read('~/GitHub/Urban_Growth_ATX/Austin_Data/Austin_MSA.geojson')%>%
  st_transform('ESRI:102741')%>%
  filter(name %in%studyAreaCounties)%>%
  st_union()

```

```{r plot_msa, warning= FALSE, message= FALSE}
ggplot() +
  geom_sf(data=austinMSA, fill = 'white') +
  geom_raster(data=rast(lc_change) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_manual(values = "#ed6728", label = "Land Cover Change")+
  labs(title = "Land Cover Change, 2009-2019", 
       fill = "") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}
lc_change2 <- lc_change

lc_change2[lc_change2 < 1] <- NA

names(lc_change2) <- "lc_change"

# ggplot() +
#   geom_sf(data=austinMSA) +
#   geom_raster(data=rast(lc_change2) %>% na.omit,
#               aes(x,y,fill=as.factor(value))) +
#   scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange") +
#   labs(title="Development Land Use Change") +
#   mapTheme
```

```{r, warning = FALSE, message = FALSE}
austinMSA_fishnet <- 
  st_make_grid(st_sf(austinMSA), 4000) %>%
  st_sf()

austinMSA_fishnet <-
  austinMSA_fishnet[austinMSA,]

# st_write(austinMSA_fishnet, "austinMSA_fishnet.shp")

```

```{r, warning = FALSE, message= FALSE}
ggplot() +
  geom_sf(data=austinMSA_fishnet) +
  labs(title="Fishnet, 4000 Foot Resolution") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}

#centroid of each raster cell transformed to a point
changePoints <-
  rasterToPoints(lc_change) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(austinMSA_fishnet))

#Join takes a long time to run - for markdown, we store the output and read it back in
# fishnet_join <- 
#   aggregate(changePoints, austinMSA_fishnet, sum) 
# 
# st_write(fishnet_join, "Austin_Data/fishnet_join.shp")
fishnet_join <- st_read("~/GitHub/Urban_Growth_ATX/Austin_Data/fishnet_join.shp")

#threshold - 100 pixels or more indicate urban growth. raster resolution is much smaller than fishnet cell size
fishnet <- fishnet_join%>%
  mutate(lc_change = ifelse(u__08_1 > 100,1,0),
         lc_change = as.factor(lc_change))%>%
  dplyr::select(lc_change)

ggplot() +
  geom_sf(data=austinMSA) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)$x, y=xyC(fishnet)$y, colour=lc_change)) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name = "") +
  labs(title = "Land Cover Development Change", subtitle = "As fishnet centroids") +
  mapTheme
```


## 2.3. Land Cover in 2009

```{r, warning = FALSE, message = FALSE}

lc_2009 <- raster("~/GitHub/Urban_Growth_ATX/Austin_Data/LC_08_rast.tif")

#read in fishnet output created from arcGIS to have FID column
# fishnet_fid<- st_read('~/GitHub/Urban_Growth_ATX/Austin_Data/Fishnet.shp')
# fishnet <- cbind(fishnet,fishnet_fid%>%st_drop_geometry)%>%
#   dplyr::select(-u__08_1)%>%
#   rename(FID = FID_1)

#Read LC_08 in, rename majority landcover type
# LC_08<- read.csv('~/GitHub/Urban_Growth_ATX/Austin_Data/LC_08.csv')%>%
#   dplyr::select(FID, MAJORITY)%>%
#   rename(LC_08 = MAJORITY)

#Read LC_19 in, rename majority landcover type
# i dont know if we need this yet
# LC_19<- read.csv('~/GitHub/Urban_Growth_ATX/Austin_Data/LC_19.csv')%>%
#   dplyr::select(FID, MAJORITY)%>%
#   rename(LC_19 = MAJORITY)
# LC_19 <- fishnet%>%left_join(., LC_19, by = 'FID')

# join landcover values from 2008 to fishnet
# fishnet <- fishnet%>%left_join(., LC_08, by = 'FID')

#Read Land Cover CSV and join to fishnet 
# Landcover graphics made in ArcGIS

ggplot() +
  geom_sf(data=austinMSA) +
  geom_raster(data=rast(lc_2009) %>% na.omit %>% filter(value > 0),
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="") +
  labs(title = "Land Cover, 2009") +
  mapTheme +
  theme(legend.direction="horizontal")
```

The table below shows the approach taken to recoded existing land cover codes into the categories used in our analysis. In the code block below new rasters are generated and `names` are applied. Naming ensures that when the raster is integrated with the fishnet, the column reflects the appropriate raster.

| Old_Classification             | New_Classification                                  |
|--------------------------------|-----------------------------------------------------|
| Open Space as well as Low, Medium and High Intensity Development | Developed |
| Deciduous, Evergreen, and Mixed Forest |  Forest |
| Pasture/Hay and Cultivated Crops | Farm |
| Woody and Emergent Herbaceous Wetlands | Woodlands |
| Barren Land, Dwarf Scrub, and Grassland/Herbaceous | Other Undeveloped |
| Water | Water |

```{r, warning = FALSE, message = FALSE}

# fishnet<- fishnet%>%
#   mutate(LC_08_reclass = case_when(LC_08 == 21 | LC_08 == 22 |LC_08 == 23 |LC_08 == 24~ "developed", 
#                            LC_08 == 41 | LC_08 == 42 |LC_08 == 43 ~ "forest",
#                            LC_08 == 81 | LC_08 == 82 ~ "farm", 
#                            LC_08 == 90 | LC_08 == 95 ~ "wetlands", 
#                            LC_08 == 52 | LC_08 == 71 |LC_08 == 31 ~ "otherUndeveloped",
#                            LC_08 == 11 ~ "water"))

developed <- lc_2009 == 21 | lc_2009 == 22 | lc_2009 == 23 | lc_2009 == 24
forest <- lc_2009 == 41 | lc_2009 == 42 | lc_2009 == 43 
farm <- lc_2009 == 81 | lc_2009 == 82 
wetlands <- lc_2009 == 90 | lc_2009 == 95 
otherUndeveloped <- lc_2009 == 52 | lc_2009 == 71 | lc_2009 == 31 
water <- lc_2009 == 11

names(developed) <- "developed"
names(forest) <- "forest"
names(farm) <- "farm"
names(wetlands) <- "wetlands"
names(otherUndeveloped) <- "otherUndeveloped"
names(water) <- "water"
```

```{r, warning = FALSE, message = FALSE}
aggregateRaster <- function(inputRasterList, theFishnet) {
  #create an empty fishnet with the same dimensions as the input fishnet
  theseFishnets <- theFishnet %>% dplyr::select()
  #for each raster in the raster list
  for (i in inputRasterList) {
  #create a variable name corresponding to the ith raster
  varName <- names(i)
  #convert raster to points as an sf
    thesePoints <-
      rasterToPoints(i) %>%
      as.data.frame() %>%
      st_as_sf(coords = c("x", "y"), crs = st_crs(theFishnet)) %>%
      filter(.[[1]] == 1)
  #aggregate to the fishnet
    thisFishnet <-
      aggregate(thesePoints, theFishnet, length) %>%
      mutate(!!varName := ifelse(is.na(.[[1]]),0,1))
  #add to the larger fishnet
    theseFishnets <- cbind(theseFishnets,thisFishnet)
  }
  #output all aggregates as one large fishnet
   return(theseFishnets)
  }
```

```{r, warning = FALSE, message = FALSE}
theRasterList <- c(developed,forest,farm,wetlands,otherUndeveloped,water)

aggregatedRasters <-
  aggregateRaster(theRasterList, austinMSA_fishnet) %>%
  dplyr::select(developed,forest,farm,wetlands,otherUndeveloped,water) %>%
  mutate_if(is.numeric,as.factor)

aggregatedRasters %>%
  gather(var,value,developed:water) %>%
  st_cast("POLYGON") %>%    #just to make sure no weird geometries slipped in
  mutate(X = xyC(.)$x,
         Y = xyC(.)$y) %>%
  ggplot() +
    geom_sf(data=austinMSA) +
    geom_point(aes(X,Y, colour=as.factor(value))) +
    facet_wrap(~var) +
    scale_colour_manual(values = palette2,
                        labels=c("Other","Land Cover"),
                        name = "") +
    labs(title = "Land Cover Types, 2009",
         subtitle = "As fishnet centroids") +
   mapTheme
```

## 2.4. Census Data

```{r, warning = FALSE, message = FALSE, results = "hide"}

census_api_key('d5e25f48aa48bf3f0766baab06d59402ea032067')

vars <- load_variables(year = 2009,
                      dataset = "acs5",
                      cache = TRUE)

AustinPop_09 <- 
  get_acs(geography = "tract", variables = "B01003_001", year = 2009,
                state = 48, geometry = TRUE, 
                county=c("Williamson", "Travis","Bastrop", "Caldwell", "Hays")) %>%
  dplyr::select( -NAME, -moe) %>%
  spread(variable, estimate)%>% 
  rename(pop_2009 = B01003_001) %>%
  st_transform(st_crs(austinMSA_fishnet))


```

```{r, warning = FALSE, message = FALSE, results = "hide"}
AustinPop_19 <- 
  get_acs(geography = "tract", variables = "B01003_001", year = 2019,
                state = 48, geometry = TRUE, 
                county=c("Williamson", "Travis","Bastrop", "Caldwell", "Hays")) %>%
  dplyr::select( -NAME, -moe) %>%
  spread(variable, estimate)%>% 
  rename(pop_2019 = B01003_001) %>%
  st_transform(st_crs(austinMSA_fishnet))
```

```{r, warning = FALSE, message = FALSE, fig.height= 8, fig.width= 11}
grid.arrange(
ggplot() +
  geom_sf(data = AustinPop_09, aes(fill=factor(ntile(pop_2009,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(AustinPop_09,"pop_2009"),
                   name="Quintile\nBreaks") +
  labs(title="Population, Houston MSA: 2009") +
  mapTheme,

ggplot() +
  geom_sf(data = AustinPop_19, aes(fill=factor(ntile(pop_2019,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(AustinPop_19,"pop_2019"),
                   name="Quintile\nBreaks") +
  labs(title="Population, Houston MSA: 2019") +
  mapTheme, ncol=2)
```

```{r, warning = FALSE, message = FALSE}
austinMSA_fishnet <-
  austinMSA_fishnet %>%
  rownames_to_column("fishnetID") %>% 
  mutate(fishnetID = as.numeric(fishnetID)) %>%
  dplyr::select(fishnetID)

fishnetPopulation09 <-
  st_interpolate_aw(AustinPop_09["pop_2009"], austinMSA_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(austinMSA_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(pop_2009 = replace_na(pop_2009,0)) %>%
  dplyr::select(pop_2009)

fishnetPopulation19 <-
  st_interpolate_aw(AustinPop_19["pop_2019"],austinMSA_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(austinMSA_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(pop_2019 = replace_na(pop_2019,0)) %>%
  dplyr::select(pop_2019)

fishnetPopulation <- 
  cbind(fishnetPopulation09,fishnetPopulation19) %>%
  dplyr::select(pop_2009,pop_2019) %>%
  mutate(pop_Change = pop_2019 - pop_2009)
```

```{r, warning = FALSE, message = FALSE, fig.height = 8, fig.width= 11}
grid.arrange(
ggplot() +
  geom_sf(data=AustinPop_19, aes(fill=factor(ntile(pop_2019,5))),colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=substr(quintileBreaks(AustinPop_19,"pop_2019"),1,4),
                   name="Quintile\nBreaks") +
  labs(title="Population, Austin MSA: 2019",
       subtitle="Represented as tracts; Boundaries omitted") +
  mapTheme,

ggplot() +
  geom_sf(data=fishnetPopulation, aes(fill=factor(ntile(pop_2019,5))),colour=NA) +
  scale_fill_manual(values = palette5,
                   labels=substr(quintileBreaks(fishnetPopulation,"pop_2019"),1,4),
                   name="Quintile\nBreaks") +
  labs(title="Population, Austin MSA: 2019",
       subtitle="Represented as fishnet gridcells; Boundaries omitted") +
  mapTheme, ncol=2)
```

```{r, warning = FALSE, message = FALSE, results = "hide"}
austinHighways <-
  st_read("~/GitHub/Urban_Growth_ATX/Austin_Data/Austin_Highways.geojson") %>%
  st_transform(st_crs(austinMSA)) %>%
  st_intersection(austinMSA)
```

```{r plot_highway, warning = FALSE, message= FALSE}
ggplot() +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2],colour=lc_change),size=1.5) +
  geom_sf(data=austinHighways, color = 'red') +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development")) +
  labs(title = "New Development and Highways",
       subtitle = "As fishnet centroids") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}

emptyRaster <- lc_change
emptyRaster[] <- NA

library(stringr)
# st_distance to highways
fishnet <- fishnet %>%
  mutate(uniqueID = as.character(row_number()))

fishnet_centroid <- fishnet %>%
  st_centroid()

highway_dist <- fishnet_centroid %>%
  st_distance(austinHighways %>%
                st_transform(st_crs(fishnet_centroid))) %>%
  as.data.frame() %>%
  mutate(uniqueID = as.character(row_number())) %>%
  gather(-uniqueID, key = "variable", value = "value") %>%
  dplyr::select(-variable) %>%
  group_by(uniqueID) %>%
 summarize(highway_dist = min(value))

highway_dist[c('num', 'str')] <- str_split_fixed(highway_dist$highway_dist, ' ', 2)
highway_dist <-highway_dist%>%
  mutate(highway_dist = as.numeric(num))%>%
  dplyr::select(uniqueID, highway_dist)


fishnet <- left_join(fishnet, highway_dist)

ggplot() +
  geom_sf(data=austinMSA) +
  geom_point(data=fishnet, aes(x=xyC(fishnet)[,1], 
                                             y=xyC(fishnet)[,2], 
                 colour=factor(ntile(highway_dist,5))),size=1.5) +
  scale_colour_manual(values = palette5,
                      # labels=quintileBreaks(fishnet,"distance_highways"),
                      name="Quintile\nBreaks") +
  geom_sf(data=austinHighways, colour = "red") +
  labs(title = "Distance to Highways",
       subtitle = "As fishnet centroids; Highways visualized in red") +
  mapTheme
```

## 2.6. The Spatial Lag of Development

```{r, warning = FALSE, message = FALSE}
nn_function <- function(measureFrom,measureTo,k) {
  #convert the sf layers to matrices
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}
```

```{r, warning = FALSE, message = FALSE}
fishnet$lagDevelopment <-
    nn_function(xyC(fishnet),
                xyC(filter(aggregatedRasters,developed==1)),
                2)

ggplot() +
  geom_sf(data=austinMSA) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2], 
                 colour=factor(ntile(lagDevelopment,5))), size=1.5) +
  scale_colour_manual(values = palette5,
                     labels=substr(quintileBreaks(fishnet,"lagDevelopment"),1,7),
                     name="Quintile\nBreaks") +
  labs(title = "Spatial Lag to 2001 Development",
       subtitle = "As fishnet centroids") +
  mapTheme
```

## 2.7. MSA Counties

```{r, warning = FALSE, message = FALSE, results = "hide"}
options(tigris_class = "sf")

studyAreaCounties <- 
  counties("Texas") %>%
  st_transform(st_crs(austinMSA)) %>%
  dplyr::select(NAME) %>%
  .[st_buffer(austinMSA,-4000), , op=st_intersects]%>%
  filter(NAME != "Burnet" & NAME != "Guadalupe")
```

```{r, warning = FALSE, message = FALSE}
ggplot() +
  geom_sf(data=studyAreaCounties) +
  geom_sf(data = austinMSA, fill = NA, color = 'red')+
  labs(title = "Study Area Counties") +
  mapTheme
```

## 2.8. Create the Final Dataset

```{r, warning = FALSE, message = FALSE}
dat <- 
  cbind(
    fishnet, fishnetPopulation, aggregatedRasters) %>%
  dplyr::select(lc_change, developed, forest, farm, wetlands, otherUndeveloped, water,
                pop_2009, pop_2019, pop_Change, highway_dist,lagDevelopment) %>%
  st_join(studyAreaCounties) %>%
  mutate(developed20 = ifelse(lc_change == 1 & developed == 1, 0, 1)) %>%
  filter(water == 0) %>%na.omit()
```

# 3. Exploratory Analysis

```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(highway_dist,lagDevelopment,lc_change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New Development as a Function of the Continuous Variables") +
    plotTheme 
```

```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(pop_2009,pop_2019,pop_Change,lc_change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New Development as a Function of Factor Variables") +
    plotTheme
```

```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(lc_change:otherUndeveloped,developed) %>%
  gather(Land_Cover_Type, Value, -lc_change, -geometry) %>%
   st_set_geometry(NULL) %>%
     group_by(lc_change, Land_Cover_Type) %>%
     summarize(n = sum(as.numeric(Value))) %>%
     ungroup() %>%
    mutate(Conversion_Rate = paste0(round(100 * n/sum(n), 2), "%")) %>%
    filter(lc_change == 1) %>%
  dplyr::select(Land_Cover_Type,Conversion_Rate) %>%
  kable() %>% kable_styling(full_width = F)
```

# 4. Predicting for 2019

## 4.2. Modeling

```{r, warning = FALSE, message = FALSE}
set.seed(3456)
trainIndex <- 
  createDataPartition(dat$developed, p = .50,
                                  list = FALSE,
                                  times = 1)
datTrain <- dat[ trainIndex,]
datTest  <- dat[-trainIndex,]

nrow(dat)

```

```{r, warning = FALSE, message = FALSE}
Model1 <- glm(lc_change ~ wetlands + forest  + farm + otherUndeveloped, 
              family="binomial"(link="logit"), data = datTrain)

Model2 <- glm(lc_change ~ wetlands + forest  + farm + otherUndeveloped + lagDevelopment, 
              family="binomial"(link="logit"), data = datTrain)
              
Model3 <- glm(lc_change ~ wetlands + forest  + farm + otherUndeveloped + lagDevelopment + pop_2009, 
              family="binomial"(link="logit"), data = datTrain)          
              
Model4 <- glm(lc_change ~ wetlands + forest  + farm + otherUndeveloped + lagDevelopment + pop_2009 + 
              pop_2019, 
              family="binomial"(link="logit"), data = datTrain)              
            
Model5 <- glm(lc_change ~ wetlands + forest  + farm + otherUndeveloped + lagDevelopment + pop_Change, 
              family="binomial"(link="logit"), data = datTrain)              
              
Model6 <- glm(lc_change ~ wetlands + forest  + farm + otherUndeveloped + lagDevelopment + pop_Change + 
              highway_dist, 
              family="binomial"(link="logit"), data = datTrain) 
```

```{r, warning = FALSE, message = FALSE}
modelList <- paste0("Model", 1:6)
map_dfc(modelList, function(x)pR2(get(x)))[4,] %>%
  setNames(paste0("Model",1:6)) %>%
  gather(Model,McFadden) %>%
  ggplot(aes(Model,McFadden)) +
    geom_bar(stat="identity") +
    labs(title= "McFadden R-Squared by Model") +
    plotTheme
```

```{r, warning = FALSE, message = FALSE}
testSetProbs <- 
  data.frame(class = datTest$lc_change,
             probs = predict(Model6, datTest, type="response")) 
  
ggplot(testSetProbs, aes(probs)) +
  geom_density(aes(fill=class), alpha=0.5) +
  scale_fill_manual(values = palette2,
                    labels=c("No Change","New Development")) +
  labs(title = "Histogram of test set predicted probabilities",
       x="Predicted Probabilities",y="Density") +
  plotTheme
```

## 4.3. Accuracy